R Notebook

### 2022-02-15 docs/3. SNA - Assignments/sna_Week3/Week3_Notebook.Rmd

# Network: Creating network Data
```{r HOLD: EDGELIST: IGRAPH: create a network object from the dataset, eval=FALSE, include=FALSE}

colnames(smugCovIdx )

# [1] "dnkName" "dnkIng"  "ingType"
# V1 <- "dnkName" 
# V2 <- "dnkIng"  
# EA <- "ingType"
dr_smugCov <- smugCovIdx

 ##If data are in edgelist format (A) and are undirected (B), use the following command:#

#igraph object 
network_ig<- graph.edgelist(as.matrix(smugCovIdx[1:2]), directed=F)

#  Adding edge and node attributes to network/graph object

#If you have edgelist data, edge attributes will often be stored in same data file
#iGraph will read in all edge attributes automatically, inspect as follows
edge_attr(network_ig)

##but to create new edge attribute if edges sorted correctly
# E(network_ig)$COLNAME<-edge.attribute$COLNAME

#The vertex names in this file must exactly match those in the network#
#sometimes this requires using as.character(x) to change from factor to character
#create vertex attributes in igraph
# V(network_ig)[match(V(network)$name,vertex_attributes$VERTEX_NAME)]$COLNAME<-
#  vertex_attributes$COLNAME

@@

```{r NETWORK EDGELIST: Statnet & Igraph, include=FALSE}
#Lets read the data into the enviroment.
#This will import it formatted as an edgelist

source_netEdge <- smugCovIdx

# Identify Edge Vertices and Edge Attributes

colnames(source_netEdge)
# [1] "dnkName" "dnkIng"  "ingType"
# V1 <- "dnkName" 
# V2 <- "dnkIng"  
# EA <- "ingType"

# Lets create an statnet object from our edgelist
#####
#First we make the object. It will have no vertex or edge attributes
sourceEL_sn <- 
  network(as.matrix(source_netEdge[1:2], 
                    matrix.type = 'edgelist')
  )

#**%e%** Let's add edge atributes, do not include the colNames for edge vertex

sourceEL_sn%e%"ingType" <- source_netEdge$ingType
sourceEL_sn%e%"dnkIng" <- source_netEdge$dnkIng

# Generating the EdgeList

##**round 1**
          # from "_netEdge" select "colNames" to be identified with "vertex.names"

temp <- source_netEdge[,c("dnkName")]

##**round 2**
            # code still selecting from "_netEdge", add "colNames"

colnames(temp)<- c("dnkIng")

##**round 3**
            # "rBind" selects "colnames" from "_netEdge" and BINDS them to temp

temp<- rbind(temp, source_netEdge[, c("dnkIng")])

##**final round** !!! is being used as variable in generating an edgelist object
                # temp contains

sourceEL_sn%v%"dnkName"<- 
  temp$dnkIng[match(sourceEL_sn%v%"vertex.names",
                    temp$dnkIng)]
      #####
            summary(sourceEL_sn)
dim(sourceEL_sn)

              # it's a bloody statnet edgelist !!!

#**igraph** version of the network
#####
          # NOTE: igraph automatically adds edge attributes, but not vertices

    #First let's create the object.

sourceEL_ig <- smugCovIdx

sourceEL_ig <- graph_from_data_frame(d = sourceEL_ig, directed = FALSE)

 #**igraph** Generate Edgelist

V(sourceEL_ig)$dnkName<-
  temp$dnkIng[
    match(V
          (sourceEL_ig)$name,
          temp$dnkIng)
  ]
# print(sourceEL_ig )
# edge_attr(sourceEL_ig)

#########
########
          smugEL_sn  <- sourceEL_sn
          smugEL_ig <- sourceEL_ig

>>>>

```{r EDGELIST: Network Objects}

#Lets make some network objects

colnames(smugCovIdx) 
# tell R data should be understood as a matrix
smug_matrix = as.matrix(smugCovIdx,
                        matrix.type="edgelist",
                        directed=FALSE) 
head.matrix(smug_matrix)

network_edgelist <- smug_matrix

  network_nodes <- smugCovIdx 
    network_nodes<- 
      select(network_nodes,
             dnkName)

    View(network_nodes)
######
      ######**igraph** 
######
        #igraph object 
network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx), 
    directed=F)
                ## create new edge attribute
  E(network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(network_ig)$ingType<-smugCovIdx$ingType
  #create vertex attributes in igraph
 V(network_ig)[match(V(smug_matrix)$dnkName,network_nodes$dnkName)]$dnkName<-
   network_nodes$dnkName

  V(network.ig)[match(V(network_edgelist)$name,network_nodes$VERTEX_NAME)]$dnkName <- network_nodes$dnkName
      summary(edge_attr(network_ig))

head(network_ig)
summary(network_ig)
      ######

#####
      #####**statnet** 
#####    
  network_edgelist <- smugCovIdx

  network_nodes <- smugCovIdx 
  network_nodes<- 
    select(network_nodes,
           dnkName)
summary(network_nodes)
View(network_nodes)

  #Now let's create a statnet object
  network_statnet <- network(network_edgelist, 
                             matrix.type = "edgelist",
                             vertex.attr = network_nodes,
                             multiple = T)

  network_statnet%e%'dnkIng'  <-
    network_edgelist$dnkIng

  network_statnet%e%'ingType'<- 
    network_edgelist$ingType

summary(network_statnet)

                                 #     access vertex attributes
  view(network_statnet%v% "vertex.names")
            # list the names of vertex attributes
  network::list.vertex.attributes(network_statnet)
            #  Generate a list the names of edge attributes
  network::list.edge.attributes(network_statnet)
            #   summarize numeric network attribute
  summary(network_statnet%v%"dnkName")

@@
